### Set-up

#### Load necessary packages

```{r}
#| label: load-packages
#| message: false
#| warning: false
library(dplyr)
library(srvyr)
library(readr)
```

## Exercises

In these exercises, you will be given a study and the data. How would you create the survey object with design variables or replicate weights, as applicable?

1.  California Health Interview Survey (CHIS) - 2023

    -   [CHIS Design](https://healthpolicy.ucla.edu/our-work/california-health-interview-survey-chis/chis-design-and-methods/chis-design)
    -   [CHIS Resources](https://healthpolicy.ucla.edu/our-work/california-health-interview-survey-chis/access-chis-data)
    -   The code below reads in a dataset select set of columns of the 2023 Adult CHIS Public Use File. Create the survey object.
    
```{r}
#| label: design-exercise-1
chis_adult_2023 <- read_rds(here::here("Exercises", "data", "chis_adult_2023.rds"))
glimpse(chis_adult_2023)
```

    
2. National Survey on Drug Use and Health (NSDUH) - 2023

    - [NSDUH Methodology](https://www.samhsa.gov/data/data-we-collect/nsduh-national-survey-drug-use-and-health/methodology)
    - [NSDUH Download Data Files](https://www.samhsa.gov/data/data-we-collect/nsduh-national-survey-drug-use-and-health/datafiles)
    - The code below reads in a dataset select set of columns of the 2023 NSDUH Public Use File. Create the survey object.
    
```{r}
#| label: design-exercise-2
nsduh_2023 <- read_rds(here::here("Exercises", "data", "nsduh_2023.rds"))
glimpse(nsduh_2023)
```

#### Bonus Exercise

   - Find a public use file of your own
   - Download the data
   - Read in the data
   - Create the survey object

