### Set-up

#### Load necessary packages

```{r}
#| label: set-up-00
#| message: false
#| warning: false
library(tidyverse)
library(gt)
```

#### Preview data

```{r}
#| label: towny-preview-00
glimpse(towny)
```

#### Exercises

1. How many different types of CSD (`csd_type`) are there in the dataset?

```{r}
#| label: exercise-00-01
```

2. How many different types of CSD and status (`status`) are there in the dataset?

```{r}
#| label: exercise-00-02
```

3. What is the proportion of each type of CSD?

```{r}
#| label: exercise-00-03
```

4. What is the proportion of each status within type of CSD?

```{r}
#| label: exercise-00-04
```

5. What is the mean population of all of the municipalities in 2021?

```{r}
#| label: exercise-00-05
```

6. What is the mean population by CSD Type in 2021?

```{r}
#| label: exercise-00-06
```

7. What is the mean population of all of the municipalities in 1996, 2001, 2006, 2011, 2016, and 2021? Try to use the across function.

```{r}
#| label: exercise-00-07
```

8. **Advanced bonus exercise** Run a t-test to see if the average population in 1996 is different from the average population in 2016.

```{r}
#| label: exercise-00-08
```
