---
title: "descriptive-exercises"
---

# Set-up

## Load necessary packages

```{r}
#| label: desc-ex-set-up
library(tidyverse)
library(srvyr)
library(srvyrexploR)
```

## Load in data and preview it

```{r}
#| label: desc-ex-data-summary
glimpse(recs_2020)
glimpse(anes_2020)
```

Find codebooks here:

  - [RECS 2020](https://tidy-survey-r.github.io/tidy-survey-book/recs-cb.html)
  - [ANES 2020](https://tidy-survey-r.github.io/tidy-survey-book/anes-cb.html)

## Create design objects for usage

```{r}
#| label: desc-ex-design-obj
anes_des <- anes_2020 %>%
  mutate(Weight = V200010b / sum(V200010b) * 231034125) %>%
  as_survey_design(
    weights = Weight,
    strata = V200010d,
    ids = V200010c,
    nest = TRUE
  )

recs_des <- recs_2020 %>%
  as_survey_rep(
    weights = NWEIGHT,
    repweights = NWEIGHT1:NWEIGHT60,
    type = "JK1",
    scale = 59 / 60,
    mse = TRUE
  )
```

# Exercises - Part 1

1. How many females have a graduate degree (according to the ANES data)?

```{r}
# Hint: The variables `Gender` and `Education` will be useful.
```

2. What percentage of people identify as “Strong Democrat”? 

```{r}
# Hint: The variable `PartyID` indicates someone’s party affiliation.
```

3. What percentage of people who voted in the 2020 election identify as “Strong Republican”? 

```{r}
# Hint: The variable `VotedPres2020` indicates whether someone voted in 2020.
```

4. What percentage of people voted in both the 2016 election and the 2020 election? Include the logit confidence interval. 

```{r}
# Hint: The variable VotedPres2016 indicates whether someone voted in 2016.
```

5. What is the design effect for the proportion of people who voted early? 

```{r}
# Hint: The variable `EarlyVote2020` indicates whether someone voted early in 2020.
```

6. What is the median temperature people set their thermostats to at night during the winter (using the RECS data)?

```{r}
# Hint: The variable `WinterTempNight` indicates the temperature that people set their thermostat to in the winter at night.
```

7. People sometimes set their temperature differently over different seasons and during the day. What median temperatures do people set their thermostats to in the summer and winter, both during the day and at night? Include confidence intervals.

```{r}
# Hint: Use the variables `WinterTempDay`, `WinterTempNight`, `SummerTempDay`, and `SummerTempNight.`
```

8. What is the correlation between the temperature that people set their temperature at during the night and during the day in the summer?

```{r}

```

9. What is the 1st, 2nd, and 3rd quartile of money spent on energy by Building America (BA) climate zone?

```{r}
# Hint: `TOTALDOL` indicates the total amount spent on all fuel, and `ClimateRegion_BA` indicates the BA climate zones.
```