---
title: "testing-exercises"
format: html
---

# Set-up

## Load necessary packages

```{r}
#| label: set-up

library(tidyverse)
library(srvyr)
library(srvyrexploR)
```

## Load in data and preview it

```{r}
#| label: data-summary
glimpse(recs_2020)
glimpse(anes_2020)
```

Find codebooks here:

  - [RECS 2020](https://tidy-survey-r.github.io/tidy-survey-book/recs-cb.html)
  - [ANES 2020](https://tidy-survey-r.github.io/tidy-survey-book/anes-cb.html)

## Create design objects for usage

```{r}
#| label: design-obj

anes_des <- anes_2020 %>%
  mutate(Weight = V200010b / sum(V200010b) * 231034125) %>%
    as_survey_design(
    weights = Weight,
    strata = V200010d,
    ids = V200010c,
    nest = TRUE
  )

recs_des <- recs_2020 %>%
  as_survey_rep(
    weights = NWEIGHT,
    repweights = NWEIGHT1:NWEIGHT60,
    type = "JK1",
    scale = 59 / 60,
    mse = TRUE
  )

```

# Exercises - Part 1

1. For households that use thermostats, they are asked what temperature they set their home during the daytime in the winter (`WinterTempDay`) and in the summer (`SummerTempDay`). Test whether daytime winter and daytime summer temperatures of homes are set the same.

```{r}
#| label: testing-exercise-1

# Hint - you'll need the recs_des object
```

2. In 2015, the average household spent $1,856 on energy expenditure^[<https://www.eia.gov/consumption/residential/data/2015/c&e/pdf/ce1.1.pdf>]. Accounting for inflation, this would be $2,036 in 2020. Test whether energy costs (`TOTALDOL`) have increased in 2020 from 2015, accounting for inflation.

```{r}
#| label: testing-exercise-2

# Hint - you'll need the recs_des object
```

# Exercises - Part 2

3. Is there a relationship between party identification (`PartyID`) and whether people voted early (`EarlyVote2020`)?

```{r}
#| label: testing-exercise-3

# Hint - you'll need the anes_des object
```

4. Is there a relationship between party identification (`PartyID`) and trust in the government (`TrustGovernment`)? Use Wald as the `statistic` option.

```{r}
#| label: testing-exercise-4

# Hint - you'll need the anes_des object
```

5. Is there a difference in the distribution of gender (`Gender`) across early voting status (`EarlyVote2020`)?


```{r}
#| label: testing-exercise-5

# Hint - you'll need the anes_des object
```

# Exercises - Part 3

6. Test whether average electric bill (`DOLLAREL`) varies by region (`Region`).

```{r}
#| label: testing-exercise-6

# Hint - you'll need the recs_des object
```
